---
title: "Assignment_7"
author: "Wiktoria Brandys"
format: html
editor: visual
---

## Code

```{r}
library(dplyr)
library(tidyr)
```

### Task 1: Load the “iris” dataset and transform it to longer

I transformed `iris` dataset to longer format using `pivot_longer command`.\
In this code I keept the `Species` and `Trait` column and assigned the values into the `Size` column.

```{r}
iris_long <- iris %>% 
  pivot_longer(cols=c(Sepal.Length, Sepal.Width,Petal.Length, Petal.Width), names_to="Trait", values_to="Size")
View(iris_long)
```

### Task 2: Answer the following questions

#### Task 2.1: What is the mean length and width of the sepal and petal of each species?

First, I grouped data by `Species` and `Trait` so than I can calculate mean size with "mean" function for each group.

```{r}
iris_long %>% 
  group_by(Species, Trait) %>% 
  summarize(mean(Size), .groups="drop") %>% 
  View()
```

#### Task 2.2: What is the largest and widest, petal and sepal, of virginica and versicolor species?

In this code, I filtered the data from column `Species` for only `virginica` and `versicolor` and than I arranged it by `Species`, `Trait` and than by `Size` in decending order.\
This arrangement is very clear for me and easy to read since frist I have all of the informations for versicolor and than for virginica.

```{r}
iris_long %>% 
  filter(Species %in% c("virginica", "versicolor")) %>% 
  arrange(Species, Trait, desc(Size)) %>% 
  View()
```

|                  | versicolor | virginica |
|------------------|------------|-----------|
| **Petal.Length** | 5.1        | 6.9       |
| **Petal.Width**  | 1.8        | 2.5       |
| **Sepal.Length** | 7.0        | 7.9       |
| **Sepal.Width**  | 3.4        | 3.8       |

#### Task 2.3: For each species, for sepal and petal width and length, how many observations are above the mean?

Just like above, I grouped the data by `Species` and `Trait` to be able to make operations only on those groups.\
Than, I created a new column named `mean_size` containing the group specific mean size.\
After that, I filtered the observations which size was above the mean group value with "\>" operator.\
Finally, I counted the number of obesrvations for each group (where the size was greater than the average).

```{r}
iris_long %>% 
  group_by(Species, Trait) %>% 
  mutate(mean_size=mean(Size)) %>% 
  filter(Size>mean_size) %>% 
  summarise(observ_above_mean = n(), .groups="drop") %>% 
  View()
```

| Species    | Trait        | observ_above_mean |
|------------|--------------|-------------------|
| setosa     | Petal.Length | 26                |
| setosa     | Petal.Width  | 16                |
| setosa     | Sepal.Length | 22                |
| setosa     | Sepal.Width  | 22                |
| versicolor | Petal.Length | 27                |
| versicolor | Petal.Width  | 22                |
| versicolor | Sepal.Length | 24                |
| versicolor | Sepal.Width  | 29                |
| virginica  | Petal.Length | 25                |
| virginica  | Petal.Width  | 23                |
| virginica  | Sepal.Length | 22                |
| virginica  | Sepal.Width  | 29                |
